--- 
title: "Disproportionate Impact of COVID-19 Disruptions on Women's Socioeconomic and Living Conditions"
author: "Xuechun Yuan, Keli Wang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

Since the outbreak of Coronavirus Disease 2019 (COVID-19), people's lives have been greatly impacted all over the world. While a substantial number of studies on the pandemic’s impacts have been conducted by researchers and delivered to the general public, a gender-based perspective is still left out of the mainstream. However, more and more reports have shown that the pandemic might have been affecting women’s living conditions disproportionately. For instance, loss of access to contraceptive measures during lockdowns might have caused a large increase in unintended pregnancies according to UN News ("Covid-19 Could Lead to Millions of Unintended Pregnancies, New Un-Backed Data Reveals"). It was also reported that women have a higher chance of being frontline workers since they make up 70% of health and social care sector workers across 104 countries (Ravanera). Therefore, in this project, we would like to take a closer look at how women’s lives have been impacted by the pandemic, and if relevant issues of gender equality have been exacerbated.

## References

"Covid-19 Could Lead to Millions of Unintended Pregnancies, New Un-Backed Data Reveals." *UN News*, United Nations, 28 Apr. 2020, https://news.un.org/en/story/2020/04/1062742. 

Ravanera, Carmina. "Primer on the Gendered Impacts of Covid-19." *Gender and the Economy*, University of Toronto, July 2021, https://www.gendereconomy.org/primer-on-the-gendered-impacts-of-covid-19/. 


<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Proposal

## Research topic

In this project, we would like to take a closer look at how women’s lives have been impacted by the pandemic, and if relevant issues of gender equality have been exacerbated. To approach this problem, we would adopt a number of different perspectives. Our first concern is the direct impact of the disease on women’s health, so we would like to investigate if women were more likely than men to be infected with, to be hospitalized for, to be admitted to ICU for, and to die of the disease; we would also like to check the gender distribution of health workers infected with or died of the disease; in addition, we would compare if women had the equal chance to be vaccinated as men. Our second concern is the influence of pandemic on women’s safety; for example, we would like to answer whether the number of domestic violence and street attacks on women have had a significant rise. Moreover, we are concerned about the pandemic’s impact on women’s socioeconomics conditions, like if women are more likely than men to be laid off, underpaid, or not paid in the context of the pandemic. If possible, we would also like to gather data on and analyze the pandemic’s impact on women’s reproductive health and access to education.

## Data availability

### WHO COVID-19 Detailed Surveillance Data Dashboard

We obtained the data from the COVID-19 Detailed Surveillance Data Dashboard of World Health Organization. WHO collected the data through official reporting to WHO, the website *Our World in Data*, the website *FIND*, and official public websites not officially reported to WHO.

The dataset contains demographic information including country, sex, and age, in addition to the daily or weekly numbers of testing, cases, hospitalizations, discharged cases, and deaths for both the general population and health workers for over 200 countries from January 2020 to the present date. Some of the data have been updated daily while the others have been updated weekly. The data is in the format of a csv file and is comprised of mostly numeric data and some categorical text data.

We would manually download the dataset in csv format from the website and import it to RSutdio.

Data Source: 
https://app.powerbi.com/view?r=eyJrIjoiYWRiZWVkNWUtNmM0Ni00MDAwLTljYWMtN2EwNTM3YjQzYmRmIiwidCI6ImY2MTBjMGI3LWJkMjQtNGIzOS04MTBiLTNkYzI4MGFmYjU5MCIsImMiOjh9

### COVID-19 Sex-Disaggregated Data Tracker

We obtained the data from the website of *The Sex, Gender and COVID-19 Project* produced collaboratively by Global Health 50/50, the African Population and Health Research Center and the International Center for Research on Women. The project has the largest database of sex-disaggregated data on COVID-19 in the world.

The dataset contains sex-disaggregated cumulative numbers and proportions of testing, cases, hospitalizations, ICU admissions, deaths, and vaccinations reported every two weeks for over 200 countries from April 2020 to June 2022.

The organizations collected sex disaggregated data on the indicators available from the WHO COVID-19 Detailed Surveillance Data Dashboard, and data for vaccinations from WHO Coronavirus Dashboard were integrated to the dashboard every month. Manual data collection was carried out every alternate month to supplement data on the indicators not available from WHO dashboard such as ICU admissions. The organizations paused manual data collection in July 2022 due to the decline in global reporting of COVID-19 data, so the final round of manual data collection was conducted in June 2022 ("Notes on the Data").

We would manually download the dataset in csv format from the website and import it to RSutdio.

Data Source: https://globalhealth5050.org/the-sex-gender-and-covid-19-project/dataset/

### Survey on Violence Against Woman During COVID-19

This set of data comes from the Rapid Gender Assessment surveys specifically focused on violence against women and COVID-19 (VAW RGAs) conducted by United Nations Women in 13 countries from April to September 2021. The data was collected by conducting face-to-face interviews along with remote surveys during the lockdowns. UN Women led this study in collaboration with Ipsos, with support from national statistical offices, national women’s machineries and a technical advisory group of experts to guide the overall survey process ("Measuring the Shadow Pandemic: Violence against Women during Covid-19" ).

Each observation in the survey dataset is a female individual’s responses to the survey questionnaire. The survey questions are composed of three parts: questions on interviewee’s demographic information, health condition and personal safety. The data is in the format of a csv file and is comprised of mostly categorical text data and some numeric data.

We would manually download the dataset in csv format from the website and import it to RSutdio.

Data Source: https://data.unwomen.org/publications/vaw-rga

### Survey on Socioeconomic Impacts of COVID-19

The data was collected by UN Women through rapid gender assessments (RGAs) conducted in at least 52 countries between March 2020 and March 2021. The assessment focused on five areas of concern: 1) economic activities and resources; 2) unpaid domestic and care work; 3) access to goods and services, 4) emotional and physical wellbeing; and 5) relief measures ("Rapid Gender Assesments on the Socioeconomic Impacts of COVID-19"). Both man and woman were interviewed in the survey, so the dataset is helpful in analyzing whether women and men are experiencing the pandemic differently.

Each observation in the survey dataset is an individual’s responses to the survey questionnaire. The survey questions are composed of questions on interviewee’s demographic information and the five areas of concern mentioned above. The data is in the format of a csv file and is comprised of mostly categorical text data and some numeric data.

We would manually download the dataset in csv format from the website and import it to RSutdio.

Data Source: https://data.unwomen.org/rga

## References 

"Measuring the Shadow Pandemic: Violence against Women during Covid-19." *UN Women Data Hub*, UN Women, 24 Nov. 2021, https://data.unwomen.org/publications/vaw-rga#_dashboardFilterRGA. 

"Notes on the Data." *Global Health 5050*, The Sex, Gender and COVID-19 Project, Apr. 2020, https://globalhealth5050.org/the-sex-gender-and-covid-19-project/notes-on-the-data/. 

"Rapid Gender Assesments on the Socioeconomic Impacts of COVID-19." *UN Women Data Hub*, UN Women, 16 Dec. 2021, https://data.unwomen.org/rga. 


<!--chapter:end:proposal.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data 

## Sources

## Cleaning / transformation

## Missing value analysis


<!--chapter:end:data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

## Summary

Females are disproportionately affected by the pandemic in many aspects. First of all, as the worldwide healthcare workforce is made up of around 70% of females, it is not surprising that a much higher number of female health workers worldwide have been infected with COVID-19 than male health workers (female:male $\approx$ 3:1); however, considerably less female health workers have died from the disease than male workers (female:male $\approx$ 2:3).

Women are more likely to be exposed to the disease but less less like to die from it. The lower fatality rate for female can be attributed to the fact that women tend to have stronger immune systems to help fight against the virus[1]. This can also be one of the reasons why woman are less likely to get hospitalized, because the might not be sick enough to go to hospital. However, it is also likely that woman are deprioritized in terms of hospitalization especially in the region where medical resources are limited (https://www.frontiersin.org/articles/10.3389/fpubh.2020.00152/full). Further study can focus on why woman are more likely to exposed to covid and whether the limited medical resources contributes to that woman are less likely to be hospitalized.

Overall, women are more negatively impacted by the pandemic in terms of economic activities, more likely to be burdened with increased domestic and care work as a result of the pandemic, and more likely to face increased difficulty in accessing goods and resources as a result of the pandemic.

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(tidyverse)
library(vcd)
library(gridGraphics)
library(gridExtra)
library(dplyr)
library(hrbrthemes)
library(countrycode)
library(forcats)
library(scales)
```

## WHO COVID-19 Detailed Surveillance Data Dashboard

```{r}
death_count <- df_hw %>%
  mutate(Count = Deaths, Type = rep("Deaths", 2)) %>%
  select(Sex, Count, Type)
  
hw_count <- df_hw %>%
  mutate(Count = Cases/1000, Type = rep("Cases\n(counted in thousands)", 2)) %>%
  select(Sex, Count, Type)

df_hw2 <- rbind(hw_count, death_count)

ggplot(df_hw2, aes(fill = Sex, y=Count, x=Sex)) +
    geom_bar(stat="identity") +
    facet_wrap(~Type, scales = "free")+
    ylab("Number of workers") +
    ggtitle("Numbers of worldwide health worker \nCOVID-19 cases and deaths") +
    scale_fill_brewer(palette = "Pastel1") +
    theme_minimal(16)
```

It can be observed from the plot that a much higher number of female health workers worldwide have been infected with COVID-19 than their male counterparts, approaching a ratio of 3:1 (female:male). This result is expected since the majority of the health workforce all over the world are female. However, even though much more female workers have gotten COVID-19 than male workers, the former are much less likely to have died of COVID-19 than the latter, as the number of worldwide health worker deaths resulted from COVID-19 approaches a ratio of 2:3 (female:male).

## COVID-19 Sex-Disaggregated Data Tracker

```{r include=FALSE}
# Clean and transform the data 
# Transform the character to date, populate variables malepop2020 and femalepop2020

covid <- covid %>%
  mutate(date = as.Date(date, format='%m/%d/%Y'), 
         vac1_date = as.Date(vac1_date, format='%m/%d/%Y'),
         vac2_date= as.Date(vac2_date, format='%m/%d/%Y'),
         cases_date= as.Date(cases_date, format='%m/%d/%Y'),
         deaths_date= as.Date(deaths_date, format='%m/%d/%Y'),
         hosp_date= as.Date(hosp_date, format='%m/%d/%Y'),
         cfr_date= as.Date(cfr_date, format='%m/%d/%Y')) %>%
  group_by(country) %>% 
  mutate(malepop2020 = first(na.omit(malepop2020)), femalepop2020 = first(na.omit(femalepop2020))) %>% ungroup()
# Add continent for each country
covid$continent <- countrycode(sourcevar = covid$country,
                            origin = "country.name",
                            destination = "continent")
covid$continent[covid$country == 'Bonaire'] <- 'Americas'
covid$continent[covid$country == 'England'] <- 'Europe'
covid$continent[covid$country == 'Kosovo'] <- 'Europe'
covid$continent[covid$country == 'Northern Ireland'] <- 'Europe'
covid$continent[covid$country == 'Saba'] <- 'Americas'
covid$continent[covid$country == 'Scotland'] <- 'Europe'
covid$continent[covid$country == 'Wales'] <- 'Europe'
```

We are interested in whether females and male are different in the probability of infection, hospitalization and death. In order to do that, we compute the rate of infection, rate of hospitalization and fatality rate for men and women respectively, and make cleveland dot plot for each rate.

```{r fig.height=20, fig.width=10}
# Prepare data to analyze the rate of infection
covid_cases <- covid %>%
  mutate(date = ifelse(is.na(cases_date) == TRUE,date,cases_date))%>%filter(is.na(cases_total_sum)==FALSE) %>% group_by(country) %>%slice_max(order_by = cases_date) %>% 
  mutate(infect_rate_male = cases_male/(malepop2020*1000),infect_rate_female = cases_female/(femalepop2020*1000))

# Cleveland dot plot
covid_cases %>%
  select(country,infect_rate_male,infect_rate_female) %>%
  gather(key='type',value='percentage',infect_rate_male,infect_rate_female) %>%
  ggplot(aes(x=percentage, y=fct_reorder2(country,type=='infect_rate_female',percentage,.desc=FALSE), color = type)) +
  geom_point() +
  ggtitle("Females Infection Rate vs. Males' Infection Rate") +
  ylab("") +
  guides(color = guide_legend(reverse = TRUE)) +
  theme(legend.position = "top")+
  theme_minimal(16)
```

```{r fig.height=20, fig.width=10}
# Prepare data to analyze hospitalization rate
covid_hosp <- covid %>%
  mutate(date = ifelse(is.na(hosp_date) == TRUE,date,hosp_date))%>%filter(is.na(hosp_total)==FALSE) %>% group_by(country) %>%slice_min(order_by = hosp_date) %>% 
  mutate(hosp_rate_male = hosp_male/(malepop2020*1000),hosp_rate_female = hosp_female/(femalepop2020*1000))
#covid_hosp

# Cleveland dot plot
covid_hosp %>%
  select('country','hosp_rate_male','hosp_rate_female') %>%
  gather(key='type',value='percentage',hosp_rate_male,hosp_rate_female) %>%
  ggplot(aes(x=percentage, y=fct_reorder2(country,type=='hosp_rate_female',percentage,.desc=FALSE), color = type)) +
  geom_point() +
  ggtitle("Females' Hospitalization Rate vs. Males' Hospitalization Rate") +
  ylab("") +
  guides(color = guide_legend(reverse = TRUE)) +
  theme(legend.position = "top")+
  theme_minimal(16)
```

```{r fig.height=20, fig.width=10}
# Prepare data to analyze fatality rate
covid_fatality <- covid %>%
  mutate(date = ifelse(is.na(deaths_date) == TRUE,date,deaths_date))%>%
  filter(is.na(deaths_total_sum)==FALSE) %>% 
  group_by(country) %>%
  slice_max(order_by = deaths_date) %>% 
  mutate(fatality_rate_male = deaths_male/cases_male,fatality_rate_female = deaths_female/cases_female)
covid_fatality <- covid_fatality %>% filter(is.na(fatality_rate_male)==FALSE & is.na(fatality_rate_female)==FALSE)

covid_fatality %>%
  select('country','fatality_rate_male','fatality_rate_female') %>%
  gather(key='type',value='percentage',fatality_rate_male,fatality_rate_female) %>%
  ggplot(aes(x=percentage, y=fct_reorder2(country,type=='fatality_rate_female',percentage,.desc=FALSE), color = type)) +
  geom_point() +
  ggtitle("Females' Fatality Rate vs. Males' Fatality Rate") +
  ylab("") +
  guides(color = guide_legend(reverse = TRUE)) +
  theme(legend.position = "top")+
  theme_minimal(16)
```

According to the plot of infection rate, females seems to more likely or equally likely to catch covid-19 in most countries. 
According to the plot of hospitalization rate, women seems to be generally less likely or equally likely to be hospitalized. 
Finally, the plot of fatality rate shows that  females are seems to be less likely to die from COVID-19 in most of countries. 

## Survey on Violence Against Woman During COVID-19

```{r}
unique(vaw$Country)
```

We first want to investigate how common is violent against woman in daily life and if the pandemic exasperate the situation. We make two bar charts and compare them. One bar chart is proportion of women who reported ever having experienced a form of VAW or knowing another woman who experienced it in lifetime for the 13 countries on the, and the other bar chart is the proportion since the pandemic.

```{r}
# barchat 1
vaw_lifetime <- vaw %>%
  group_by(Country, rC30_grouped) %>%
  summarise(cnt = n()) %>%
  mutate(prop_vaw_lifetime = round(cnt / sum(cnt), 3)) %>% 
  arrange(desc(prop_vaw_lifetime))  %>% filter(rC30_grouped=='Yes')
#vaw_lifetime

ggplot(data=vaw_lifetime, aes(x=fct_reorder(Country, prop_vaw_lifetime, .desc = T), y=prop_vaw_lifetime*100)) +
geom_bar(stat="identity",  fill='orange2') +
  ylim(0, 100)+
  geom_text(aes(label= prop_vaw_lifetime*100), vjust=-1) +
  theme(axis.text.x = element_text(angle = 40), axis.title.x = element_blank())+
  labs(y= "Percentage", title = 'Proportion of women who reported ever having experienced a form of VAW or \n knowing another woman who experienced a form of violence')+theme_minimal(16)

# barchart2
vaw_since_covid_cnt <- vaw %>% filter(C30_1_recode =='Yes'| C30_2_recode=='Yes' | C30_3_recode=='Yes'| C30_4_recode =='Yes' | C30_5_recode=='Yes'|C30_1_recode =='Yes, since the start of COVID-19'| C30_2_recode=='Yes, since the start of COVID-19' | C30_3_recode=='Yes, since the start of COVID-19'| C30_4_recode =='Yes, since the start of COVID-19' | C30_5_recode=='Yes, since the start of COVID-19') %>% 
  group_by(Country) %>%  
  summarise(cnt = n())%>%
  arrange(desc(cnt))%>%ungroup()
#vaw_since_covid_cnt

vaw_total_cnt <- vaw %>% group_by(Country) %>% summarise(total_cnt = n()) %>% ungroup() %>% select(Country, total_cnt)
prop_vaw_aft_covid <- left_join(vaw_since_covid_cnt, vaw_total_cnt, by = c('Country'='Country')) %>% mutate(prop_vaw_aft_covid = round(cnt/total_cnt,3))
#prop_vaw_aft_covid

ggplot(data=prop_vaw_aft_covid, aes(x=fct_reorder(Country, prop_vaw_aft_covid, .desc = T), y=prop_vaw_aft_covid*100)) +
geom_bar(stat="identity",  fill='orange2') +
  ylim(0, 100)+
  geom_text(aes(label= prop_vaw_aft_covid*100), vjust=-1) +
  theme(axis.text.x = element_text(angle = 40), axis.title.x = element_blank())+
  labs(y= "Percentage", title = 'Proportion of women who reported ever having experienced a form of VAW or \n knowing another woman who experienced a form of violence \n since the COVID-19 pandemic')+theme_minimal(16)
```

From the first bar chat we can observe that violence against women is a pre-existing problem. In 9 out of 13 countries, more than a half of women have ever experienced or a form of violence or know someone who experienced in lifetime. The proportion since the pandemic is also high. For all countries, more than one forth of women have experienced violence, which is definitely a high proportion given the pandemic has lasted for only three years. Clearly, The pandemic exasperates the violence against women.

Then we want to investigate the population that are most susceptible to violence against women. To do that, we make 3 mosaic plots to investigate the correlation between the proportion of women who experienced VAW (violence against women) and the following four aspects: age, locality, living with children or not  and employment status. 

```{r}
# Fix the level labels in some columns
vaw_recode <- vaw %>% mutate(C30_1_recode = recode_factor(vaw$C30_1_recode, Refusal = "other", Refused = "other", "Don't know" = 'other', "Yes, since the start of COVID-19" = 'Yes'),
               C30_2_recode = recode_factor(vaw$C30_1_recode, Refusal = "other", Refused = "other", "Don't know" = 'other', "Yes, since the start of COVID-19" = 'Yes'),
               C30_3_recode = recode_factor(vaw$C30_1_recode, Refusal = "other", Refused = "other", "Don't know" = 'other', "Yes, since the start of COVID-19" = 'Yes'),
               C30_4_recode = recode_factor(vaw$C30_1_recode, Refusal = "other", Refused = "other", "Don't know" = 'other', "Yes, since the start of COVID-19" = 'Yes'),
               C30_5_recode = recode_factor(vaw$C30_1_recode, Refusal = "other", Refused = "other", "Don't know" = 'other', "Yes, since the start of COVID-19" = 'Yes')) %>% 
  mutate(if_vaw_aft_covid = ifelse(C30_1_recode =='Yes'| C30_2_recode=='Yes' | C30_3_recode=='Yes'| C30_4_recode =='Yes' | C30_5_recode=='Yes', 'Yes','No')) 
#unique(vaw_recode$if_vaw_aft_covid)
```
```{r}
par(mfrow = c(1, 3)) 
vcd::mosaic(if_vaw_aft_covid ~ AgeCat, vaw_recode, direction = c("v", "h"), 
            set_labels=list(AgeCat = c('18-29','30-39','40-49','50-59','>=60')),
            set_varnames = c(if_vaw_aft_covid="If reported VAW since the pandemic", AgeCat="Age"),
            margins = c(3, 3),
            las=2,
            highlighting_fill= c("#cc9966","#ff99ff"),
            rot_labels=c(0,0))

vcd::mosaic(if_vaw_aft_covid ~ child, vaw_recode, direction = c("v", "h"), 
            set_varnames = c(if_vaw_aft_covid="If reported VAW since the pandemic", child="living with children"),
            margins = c(3, 3),
            las=2,
            highlighting_fill= c("#cc9966","#ff99ff"),
            rot_labels=c(0,0))

vcd::mosaic(if_vaw_aft_covid ~ Locality, vaw_recode, direction = c("v", "h"), 
            set_labels=list( Locality = c('rural', 'urban')),
            set_varnames = c(if_vaw_aft_covid="If reported VAW since the pandemic"),
            margins = c(3, 3),
            las=2,
            highlighting_fill= c("#cc9966","#ff99ff"),
            rot_labels=c(0,0))

vaw_employment <- vaw_recode %>% filter(is.na(BR_rA03) == FALSE)
vcd::mosaic(if_vaw_aft_covid ~ BR_rA03, vaw_recode, direction = c("v", "h"), 
            set_labels=list( Locality = c('rural', 'urban')),
            set_varnames = c(if_vaw_aft_covid="If reported VAW since the pandemic", BR_rA03="Employment"),
            margins = c(3, 3),
            las=2,
            highlighting_fill= c("#cc9966","#ff99ff"),
            rot_labels=c(40,0))
```

Woman aged between 18 and 29 are most likely to reported their experiences of VAW since pandemic and women older than 60 are least like to do that. It is possible that young women are more likely to be the victim and also they are more willing to tell others about the experience and seek help.   

Those who live with children are more likely to experience violence. One possible reason is that those who have children usually resident with their partner or husband, and thus more likely to be victims of domestic violence. 

Woman living in rural area are more likely to experience violence. This is probably because rural areas are less populated and the predators tend to commit crime in such areas.

Finally, women that are unemployed are much more likely to be treated inapproriately. They may economically dependent on their spouse or partner, and thus tend to endure domestic violence.

```{r}
vaw_locality <- vaw_recode %>% filter((rS11 == 'Rural' | rS11 == 'Urban') & (C03 == "Safe"|C03 == "Very safe"|C03 == "Not safe at all"|C03 == "Not very safe"))%>% select(rS11, C03) %>% rename(locality = rS11, level_feeling_safety = C03)
#vaw_locality
vaw_locality_summary <- vaw_locality %>% 
  group_by(locality, level_feeling_safety) %>% 
  count(name = "n_answers") %>% 
  group_by(locality) %>% 
  mutate(percent_answers = n_answers / sum(n_answers)) %>% 
  ungroup() %>% 
  mutate(percent_answers_label = percent(percent_answers, accuracy = 1))
#vaw_locality_summary
vaw_locality_summary_diverging <- vaw_locality_summary %>%
  mutate(percent_answers = if_else(level_feeling_safety %in% c("Very safe", "Safe"), percent_answers, -percent_answers)) %>% 
  mutate(percent_answers_label = percent(percent_answers, accuracy = 1))
#vaw_locality_summary_diverging

vaw_locality_summary_diverging_right_order %>%
  ggplot(aes(x = locality, 
             y = percent_answers,
             fill = level_feeling_safety)) +
  geom_col() +
  geom_text(aes(label = percent_answers_label),
            position = position_stack(vjust = 0.5),
            color = "white",
            fontface = "bold") +
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(breaks = "Not safe at all", "Not very safe", "Safe", "Very safe",
                    values = c(
                      "Not safe at all" = "darkorange3",
                      "Not very safe" = "orange",
                      "Safe" = "deepskyblue",
                      "Very safe" = "deepskyblue4"
                    )) +
  labs(title = "How safe did respondent feel to walk alone at night?",
       x = NULL,
       fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = "top")+theme_minimal(16)
```

To confirm our proposed reason why women living in rural area are more likely to experienced violence, we made a diverging plot on levels of safety that the respondents feels when they walk alone at night. It shows that slightly more proportion of women living n rural feel unsafe than those living in urban areas.

## Survey on Socioeconomic Impacts of COVID-19

### Impact on economic activities

```{r fig.height=6, fig.width=6, warning=FALSE, message=FALSE}
count1 <- df %>%
  mutate(lost_job = fct_rev(lost_job)) %>%
  group_by(lost_job, sex) %>%
  summarize(Freq = n()) %>%
  filter(lost_job !="" & sex !="")

count2 <- df %>%
  mutate(insurance_unemp = fct_rev(insurance_unemp)) %>%
  group_by(insurance_unemp, sex) %>%
  summarize(Freq = n()) %>%
  filter(insurance_unemp !="" & sex !="")

count3 <- df %>%
  mutate(work_hours = fct_rev(work_hours)) %>%
  group_by(work_hours, sex) %>%
  summarize(Freq = n()) %>%
  filter(work_hours !="" & sex !="")

count4 <- df %>%
  mutate(income_loss = fct_rev(income_loss)) %>%
  group_by(income_loss, sex) %>%
  summarize(Freq = n()) %>%
  filter(income_loss !="" & sex !="")


grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))

pushViewport(viewport(layout.pos.col = 1, layout.pos.row = 1))
mosaic(lost_job ~ sex, count1,
            direction = c("v", "h"),
            highlighting_fill = RColorBrewer::brewer.pal(2, "Accent"),
            labeling = labeling_border(set_varnames = c(lost_job = "Lost Job",
                                                        sex = "Sex")),
            labeling_args=list(gp_labels=gpar(fontsize=16),
            gp_varnames=gpar(fontsize=16)),
            newpage = FALSE)
popViewport()

pushViewport(viewport(layout.pos.row = 1, layout.pos.col = 2))
mosaic(insurance_unemp ~ sex, count2,
            direction = c("v", "h"),
            highlighting_fill = RColorBrewer::brewer.pal(2, "Accent"),
            labeling = labeling_border(set_varnames = c(insurance_unemp = "Unemployment insurance",
                                                        sex = "Sex")),
            labeling_args=list(gp_labels=gpar(fontsize=16),
            gp_varnames=gpar(fontsize=16)),
            newpage = FALSE)
popViewport()

pushViewport(viewport(layout.pos.row = 2, layout.pos.col = 1))
mosaic(work_hours ~ sex, count3,
            direction = c("v", "h"),
            highlighting_fill = RColorBrewer::brewer.pal(4, "Accent"),
            labeling = labeling_border(set_varnames = c(work_hours = "Work Hours Change",
                                                        sex = "Sex"),
                                       abbreviate_labs = c(FALSE, 3, 6)),
            labeling_args=list(gp_labels=gpar(fontsize=16),
            gp_varnames=gpar(fontsize=16)),
            newpage = FALSE)
popViewport()

pushViewport(viewport(layout.pos.row = 2, layout.pos.col = 2))
mosaic(income_loss ~ sex, count4,
            direction = c("v", "h"),
            highlighting_fill = RColorBrewer::brewer.pal(2, "Accent"),
            labeling = labeling_border(set_varnames = c(income_loss = "Lost Income",
                                                        sex = "Sex")),
            labeling_args=list(gp_labels=gpar(fontsize=16),
            gp_varnames=gpar(fontsize=16)),
            newpage = FALSE)
popViewport()

```

It can be observed that compared to men, women are more likely to lose their jobs and suffer from increased work hours after the pandemic. However, it is shown that women are also more likely to be covered by unemployment insurance than men, and that men are slightly more likely to suffer from income loss than women as a result of the pandemic.

### Impact on unpaid domestic and care work

```{r}
hh_ct = table(filter(df[c(2,7)], hh_resp == "yes"))
child_ct = table(filter(df[c(2,8)], child_rep == "yes"))
adult_ct = table(filter(df[c(2,9)], adult_care == "yes"))
p_ct = table(filter(df[c(2,10)], partner_support == "yes"))

n_women = sum(df$sex == "Women")
n_men = sum(df$sex == "Men")
count = c(hh_ct, child_ct, adult_ct, p_ct)
total = rep(c(n_men, n_women), 2)
percentage = count/total

sex = rep(c("Men", "Women"), 4)
activity = c(rep("domestic", 2), rep("childcare", 2), rep("adultcare", 2), rep("partner_support", 2))
df_hc = data.frame(activity, sex, percentage)

ggplot(df_hc, aes(fill=sex, y=percentage, x=activity)) + 
    geom_bar(position="dodge", stat="identity") +
    ggtitle("Percentage of respondents burdened with \nincreased domestic and care work") +
    xlab("Domestic and care work type") +
    scale_fill_brewer(palette = "Pastel2") +
    theme_minimal(16)
```

It can be observed that women are more likely than men to be burdened with at least one increased domestic activity, at least one increased adultcare activity, and at least one increased childcare activity. It is also worth noting that, on the contrary, men are more likely than women to have received support from their spouse for domestic and care activities.

### Impact on access to goods and resources

```{r}
food = table(filter(df[c(2,11)], food_access == "yes"))
prod = table(filter(df[c(2,12)], access_product == "yes"))
care = table(filter(df[c(2,13)], access_care == "yes"))
fr_sup = table(filter(df[c(2,14)], friends_support == "yes"))
n_women = sum(df$sex == "Women")
n_men = sum(df$sex == "Men")

count = c(food, prod, care, fr_sup)
total = rep(c(n_men, n_women), 2)
percentage = count/total
sex = rep(c("Men", "Women"), 4)
resource = c(rep("food", 2), rep("sanitary_products", 2), 
             rep("medical_care", 2), rep("friends_support", 2))
ordered = c("friends_support", "medical_care", "sanitary_products", "food")
resource = factor(resource, levels = ordered)
df_res = data.frame(resource, sex, percentage)


ggplot(df_res, aes(sex, resource, fill= percentage)) +
  geom_tile() +
  scale_fill_gradient(low="white", high="blue") +
  ggtitle("Percentage of respondents with\n increased difficulty in \naccessing goods & resources") +
  ylab("goods/resource type") +
  theme_ipsum() +
  theme_minimal(16)
```

It can be observed that after the pandemic, compared to men, women are more likely to face increased difficulty in accessing food, sanitary products, and medical care (women have darker/bluer shades in these categories in the plot). Men and women are about equally likely to get support from relatives and friends.

<!--chapter:end:results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component

We made three horizontal bar charts to analyse the impact of pandemic including change in women's working hours, job loss and income loss. You may play with the interactive plot here


<script src="https://d3js.org/d3.v7.js"></script>
<script>
    <!-- Initialize a select button -->
    <select id="selectButton"></select>
    <!-- Color Scale -->
    <script src="https://d3js.org/d3-scale-chromatic.v1.min.js"></script>
    <div id="my_dataviz"></div>
</script>  

<script>
// Create svg and initial bars

  const w = 460;
  const h = 340;
  const margin = {top: 20, right: 30, bottom: 40,
      left: 150};
  const innerWidth = w - margin.left - margin.right;
  const innerHeight = h - margin.top - margin.bottom;
  
  
  var groups = ['Reported job loss of Women', 'Reported income loss of Women', 'Impact on hours worked of Women']
   // add the options to the button
  d3.select("#selectButton")
      .selectAll('myOptions')
     	.data(groups)
      .enter()
    	.append('option')
      .text(function (d) { return d; }) // text showed in the menu
      .attr("value", function (d) { return d; }) // corresponding value returned by the button
  
  var myColor = d3.scaleOrdinal()
      .domain(groups)
      .range(d3.schemeSet2);
  
  const bardata1 = [{region: "Americas and the Caribbean", value: 49},
                 {region: "Asian and the Pacificb", value: 28},
                 {region: "All regions", value: 23},
                 {region: "Esatern and Southern Africa", value: 19},
                 {region: "Europe and Central Asia", value: 14},
                 {region: "Arab States", value: 11},
                 {region: "West and Central Africa", value: 8}];
                 
  const bardata2 = [{region: "West and Central Africa", value: 87},
                 {region: "Americas and the Caribbean", value: 73},
                 {region: "All regions", value: 61},
                 {region: "Esatern and Southern Africa", value: 57},
                 {region: "Arab States", value: 54},
                 {region: "Europe and Central Asia", value: 39},
                 ];
                 
  const bardata3 = [{region: "Decreased", value: 48},
                 {region: "No change", value: 40},
                 {region: "Increased", value: 13}];
                 
  bardata = bardata1
      
  const yScale = d3.scaleBand()
      .domain(bardata.map(d => d.region))
      .range([0, innerWidth])
      .paddingInner(.1);
          
    
  const xScale = d3.scaleLinear()
          .domain([0, d3.max(bardata.map(d => d.value))])
          .range([0,innerWidth]);
    
  const xAxis = d3.axisBottom()
          .scale(xScale);
    
  const yAxis = d3.axisLeft()
        .scale(yScale);
    
  // add svg
    
  const svg = d3.select("body")
        .append("svg")
          .attr("width", w)
          .attr("height", h);
    
    // add background rectangle
    
  svg.append("rect")
          .attr("x", 0)
          .attr("y", 0)
          .attr("width", w)
          .attr("height", h)
          .attr("fill", "aliceblue");
    
    // add bars as a group
    
  const bars = svg.append("g")
  .attr("id", "plot")
  .attr("transform", `translate (${margin.left}, ${margin.top})`)
  .selectAll("rect")
  .data(bardata);
    
  bars.enter().append("rect").transition().duration(600)
  .attr("y", d => yScale(d.region))
  .attr("x", d => xScale(0))
  .attr("height", yScale.bandwidth())
  .attr("width", d => xScale(d.value))
  .attr("fill", 'orange');
    
  // add axes
    
  svg.append("g")
  .attr("class", "xAxis")
  .attr("transform", `translate (${margin.left}, ${h - margin.bottom})`)
  .call(xAxis);
    
  svg.append("g")
  .attr("class", "yAxis")
  .attr("transform", `translate (${margin.left}, ${margin.top})`)
  .call(yAxis);
  
  function update(selectedGroup) {
    if (selectedGroup == groups[0]) {
        bardata = bardata1; color = 'orange';}else if(selectedGroup == groups[1]){
        bardata = bardata2; color = 'green';}else{
        bardata = bardata3; color = 'blue';}
    const yScale = d3.scaleBand()
          .domain(bardata.map(d => d.region))
          .range([0, innerWidth])
          .paddingInner(.1);
    const xScale = d3.scaleLinear()
          .domain([0, d3.max(bardata.map(d => d.value))])
          .range([0,innerWidth]);
    const xAxis = d3.axisBottom()
          .scale(xScale);
    const yAxis = d3.axisLeft()
          .scale(yScale);
    
    // add svg
    
    
 // add background rectangle
    svg.append("rect")
          .attr("x", 0)
          .attr("y", 0)
          .attr("width", w)
          .attr("height", h)
          .attr("fill", "aliceblue");
  // add bars as a group
    const bars = svg.append("g")
          .attr("id", "plot")
          .attr("transform", `translate (${margin.left}, ${margin.top})`)
        .selectAll("rect")
          .data(bardata);
    bars.enter().append("rect").transition().duration(600)
          .attr("y", d => yScale(d.region))
          .attr("x", d => xScale(0))
          .attr("height", yScale.bandwidth())
          .attr("width", d => xScale(d.value))
          .attr("fill", color);
    svg.append("g")
          .attr("class", "xAxis")
          .attr("transform", `translate (${margin.left}, ${h - margin.bottom})`)
          .call(xAxis);
    svg.append("g")
          .attr("class", "yAxis")
          .attr("transform", `translate (${margin.left}, ${margin.top})`)
          .call(yAxis);
  }
        
  // When the button is changed, run the updateChart function
  d3.select("#selectButton").on("change", function(d){
        // recover the option that has been chosen
        var selectedOption = d3.select(this).property("value");
        // run the updateChart function with this selected option
        update(selectedOption, svg);
    })

</script>



<!--chapter:end:interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion

It can be concluded that females are more negatively affected by the pandemic than males in many aspects. The 70% of females in the global healthcare workforce have borne the brunt of the pandemic, as a much higher number of female health workers worldwide have been infected with COVID-19 than male health workers (female:male $\approx$ 3:1); namely, 75% of the infected health workers are females, which is greater than the percentage of 70% in the global healthcare workforce, so woman workers are also more likely than man workers to be infected with COVID-19. However, considerably less female health workers have died from the disease than male workers despite much more female workers were infected than male workers (female:male $\approx$ 2:3), which may be due to women's stronger immune system.

Overall, women's socioeconomic conditions are also more negatively affected by the pandemic than men's. Firstly, women are more negatively impacted by the pandemic in terms of economic activities, since they are more likely to lose their jobs and suffer from increased work hours after the pandemic, although they are also more likely to be covered by unemployment insurance and slightly less likely to suffer from income loss than men as a result of the pandemic. Secondly, women are more likely to be burdened with increased domestic and care work as a result of the pandemic, as they are more likely than men to be burdened with increased domestic activity, increased adultcare activity, and increased childcare activity, while being less likely than men to have received support from their spouse for domestic and care activities. Thirdly, women are more likely than men to face increased difficulty in accessing food, sanitary products, and medical care as a result of the pandemic.

One limitation of our project is that we are not able to explore the data in every dimension possible; for example, numbers of female and male health worker COVID-19 cases and deaths were only explored on the worldwide level and not by continents or countries. A main reason is that the data for some continents are of single digits and therefore is not meaningful for discovering trends across continents. A future direction could be to select some countries with large enough numbers that could produce meaningful results and explore trends across countries. Another limitation is that although the 5 datasets for the socioeconomic impacts of COVID-19 have been provided for 5 regions/continents and at least 52 countries by UN Women's RGA, they were mostly explored in the form of a combined dataset and therefore many trends across the regions/continents or countries were not spotted. Hence, another future direction could be exploring these trends across the regions/continents and/or countries.

From our project, it can be concluded that females are more negatively affected by the pandemic than males in many aspects. More women are in the healthcare workforce and women are also more likely to be infected with COVID-19 than men, but the media scarcely report or praise female workers' sacrifice and contribution, and some media even have portrayed female workers as unwilling to contribute. Women are more likely to suffer from increased violence, worsened socioeconomic conditions, and greater lack of access to resources than men, but far less than enough public attention or effort is directed to their disproportionate hardship caused by the pandemic. Therefore, awareness should be raised and more efforts should be made toward closing these various gender gaps that have been widened by the pandemic.

<!--chapter:end:conclusion.Rmd-->

